---
title: "Mapping Chinese exports within the BRICS in from 2013 to 2024"
author: "Vinicius Paulinelli"
date: "2025-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This project explores China's commercial relations with other members of the BRICS.

The data set is from UNCOMTRADE and focuses on Chinese exports from 2013 to 2024, when the group was expanded to include Ethiopia, Iran, Egypt and the UAE. Indonesia became a full member in early 2025.

The main hypothesis are:
1. Intra-bloc trade has been increasing due to political proximity;
2. An increase in exports might be an early indicator of membership. In other words, we expect China to increase its exports to other countries in the years preceding their entry.

```` {r setting wd and csv files, echo = FALSE, include = FALSE}

library(ggplot2)
library(tidyverse)
library(readr)
setwd('C:/Users/Vinicius/Desktop/Pessoal/Data Projects/BRICS trade')
TradeData_10_30_2025_11_17_21 <- read_csv("TradeData_10_30_2025_11_17_21.csv")

````

## Getting and Transforming Data

One can download the file directly from UNCOMTRADE's platform with the desired parameters (country of focus, destinations, time period and etc). I just needed to filter the raw data to contain only the relevant parts and changed some names for clarity.

Since the amount was in total USD, I rounded up the unities to billions to facilitate calculations.

```{r getting data}
#loading and transforming the dataset
trade_data_2 <- select(TradeData_10_30_2025_11_17_21, 
                       refYear, 
                       partnerDesc,
                       primaryValue) %>%
  rename('Year' = refYear,
         'Country' = partnerDesc,
         'Total'= primaryValue)

#converting to billions of USD
trade_data_round_2 <- mutate(trade_data_2, Total = round(Total / 1000000000, 3))
head(trade_data_round_2)

```

## Exploring the Data

### What was the temporal behavior of exports?

Trade remained stable until 2020 but has been increasing ever since, even beyond pre-pandemic levels. So we have a positive evidence for our first hypothesis.

```{r total exports, echo=TRUE}
total_yearly_trade <- trade_data_round_2 %>%
  group_by(Year) %>%
  summarise(total_year = sum(Total))

ggplot(total_yearly_trade, aes(as.factor(Year), total_year)) + 
  geom_col(fill = 'steelblue', colour = 'black') +
  labs(title = 'Total Chinese Exports from 2013 to 2024',
       x = 'Year',
       y = 'Total trade (in billions of USD)')
```

### How are the exports divided?

We can also get a glimpse of which countries were China's most valuable in the bloc, if we vizualize how exports are divided.


```{r country percentage, echo=TRUE}
#Getting the total value by Country
total_trade_country <- trade_data_round_2 %>% 
  group_by(Country) %>%
  summarise(total_country = sum(Total))


ggplot(total_trade_country, aes(total_country, Country)) + 
  geom_col(fill = 'steelblue', colour = 'black') + 
  labs(title = 'Total value of exports by Country (2013-2024)',
       x = 'Total (in billions of USD)',
       y = 'Country')

```

Looks like India, Russia and Indonesia were the top three - USD 958.548 bn, USD 735.944 bn, and USD 580.048 bn, respectively..

It would be interesting to break down these results by type of good and service in the future. I would assume they differ a lot.

### How were the new members included?

And finally we can wrap it all together and analyze exports over time focusing only on thr new members added in 2024-25

```{r new members, echo=TRUE}
#Mapping trade with the newest members
total_new_members <- trade_data_round_2  %>%
  filter(Country %in% c('Egypt','Ethiopia','Iran','Indonesia','United Arab Emirates'))

ggplot(total_new_members, aes(as.factor(Year), Total, fill = Country)) + 
  geom_col(position = 'dodge', colour = 'black') +
  theme(legend.position = 'bottom') + 
  labs(title = 'Total trade value by country (2013-2024)',
       x = 'Year',
       y = 'Total (in billions of USD)')
```

The most interesting trends to observe:

1. For Indonesia and the UAE, exports were already increasing in the years prior to their adhesion. This might indicate that China's support for their adhesion was linked to securing a relevant consumer market.

2. The same is not true for Egypt, Ethiopia and Iran. Their adhesion might have been more due to political and/or geopolitical considerations than pure economic reasons.

The conclusion is two-fold: From China's perspective, the new BRICS is about a mix of commercial and political advantages to its standing in global affairs. 


